### 2019-09-06 manuscript/manuscript.tex
ms edits, plots

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change ($\Delta_{_{10}} p_t$) in some initial reference generation $s$ (the row
of the matrix), and some later timepoint $t$ (the column of the matrix). We
corrected these matrices for biases created due to sampling noise, and
normalize the entries for heterozygosity (see Supplementary Material Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
the bias correction). The covariances are expected to be zero when only drift
is acting in a closed population, as only heritable variation for fitness can
create covariance between allele frequency changes \parencite{Buffalo2019-io}.
Averaging across the ten replicate temporal covariances matrices, we find
temporal covariances that are statistically significant (95\% block bootstraps
CIs do not contain zero), consistent with linked selection perturbing
genome-wide allele frequency changes over very short time periods. The
covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals, as expected when
directional selection affects linked variants' frequency trajectories until
ultimately linkage disequilibrium and additive genetic variance for fitness
decay \parencite{Buffalo2019-io}. The temporal covariances per replicate are
noisier but this general pattern holds; see Supplementary Figures XXX.
\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
time intervals are on average ten generations apart, and given the temporal
decay in covariance we see, the covariances on shorter time-scales may well be
higher yet (see Supplementary Material Section \ref{supp:barghi-covs} for more
details).

@@

Finally, we observed that in the \gc{longest study}  we   %run evolve-and-resequence 
analyzed \parencite{Barghi2019-qy}, some genome-wide temporal covariances
\gc{become} negative at future timepoints (see the first two rows in
Figure \ref{fig:figure-1} A). \gc{This shows that alleles that were on
average going up initially are no going down in frequency, i.e. that
the average direction of selection experienced by alleles has
flipped. This must reflect either a change in the environment or the
genetic background, due to changing epistatic relationships among
alleles or  recombination breaking up selective alleles.}
\gc{Such reversals of selective dynamics could be occuring at other time-points but} the
signal of a change in the direction of selection \gc{at particular
  loci} may be washed out when we calculate our  genome-wide \gc{average} temporal
covariances. To address this limitation, we
calculated \gc{the distribution of the} temporal covariances over
100kb genomic windows \gc{(Figure \ref{fig:figure-3}, pooling
across all replicates see Supplementary Figure \ref{suppfig:barghi-offset-replicate-panels} for individuals replicates).}
\gc{The}
covariance of each tile is noisy, \gc{due to sampling and genetic
  drift, shown in }. \gc{The neutral distribution of the covariance is
  complicated due to} linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). 
\gc{Thus} we have developed a permutation-based procedure that constructs a null
distribution by randomly flipping the signs of the    %of the
                                %genome-wide covariances as they would
                                %look if only neutral genetic drift was acting. We do this 
allele frequency changes per-genomic window. \gc{This destroys}  %%  GC bit about 1,000 times seems more like figgure caption or methods
the systematic covariances created by linked selection but creates a sampling
distribution of the covariances spuriously created by neutral genetic
drift while preserve the complex dependencies
between adjacent loci created by linkage disequilibrium (see Supplementary
Material Section \ref{supp:empirical-null} for more details on this approach).
This empirical neutral null distribution is conservative in the sense that the
variances of the covariances are wider than expected under drift
alone \gc{as they include the effect of selection on the alele
  frequency change within a time-interval, just not between
  time-intervals.} We see, Figure \ref{fig:figure-3} A and B, that
windowed temporal covariances between close timepoints are skewed positive (a
heavy right tail), while between more distant timepoints these windowed
temporal covariances tend to shift to become more negative (a heavy
left tail).
We quantified the degree to which the left and right
tails \gc{are inflated compared to the null distribution as a function
  of time}, and see excesses in both tails in Figure
\ref{fig:figure-3} C. This finding is also
robust to sign-permuting allele frequency changes on a chromosome-level, the
longest extent that gametic linkage disequilibria can extend {\bf ????} (Supplementary
Figure \ref{suppfig:barghi-tailprobs-seqid}).  \gc{We} see a striking
pattern that \gc{the windowed} covariances not only decay towards zero, but in fact become
negative through time, consistent with \gc{many} regions in the genome having had a
reversed fitness effect at later timepoints.

@@

Doc Edge.

@@

@@

We also applied our temporal covariance approach to \textcite{Bergland2014-ij},
which found evidence of genome-wide fluctuating selection between Spring and
Fall seasons across three years \emph{Drosophila melanogaster}. As described in
\textcite{Buffalo2019-io}, we might expect positive covariances between like
seasons (i.e. Spring 2010 and Spring 2011), and negative covariances between
dislike seasons (i.e. Spring 2010 and Fall 2011) when fluctuating selection is
strong and acts genome-wide, as the original study found. However, while we
find temporal covariances that are non-zero, we find only weak support for a
seasonal fluctuating model driving these covariances. In Supplementary Figure
XXX, we show the temporal covariances from varying reference generations,
across Spring/Fall, Spring/Spring, and Fall/Fall pairs. The first row of
temporal covariance matrix is consistent with fluctuating selection operating
for two timepoints, as the first covariance is negative, and the second is
positive, and later covariances are not statistically differentiable from zero
(which could occur if LD and additive genetic variance decay). However, the all
other temporal covariances do not fit the pattern we would expect under
genome-wide fluctuating selection.

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[]{figures/bergland-pvalue-permutation-hist.pdf}
  \caption{}
  \label{suppfig:bergland-pvalue-hist}
\end{figure}

@@

To investigate whether there is a genome-wide evidence of an enrichment of
fluctuating selection we created an empirical null distribution by randomly
permuting the season labels and re-running the per-locus seasonal GLM model, as
proposed by \parencite{Machado2018-cs}. We find, regardless of whether we
permute at the locus-level or the permutation replicate-level, that the
observed seasonal p-value distribution \textcite{Bergland2014-ij} is not
enriched for significant p-values beyond what we would expect from the
permutation null. In fact, there appears there is more enrichment for low
p-values when seasonal labels are randomly permuted (Supplementary Figure
\ref{suppfig:bergland-pvalue-hist}, suggesting by random chance we might expect
more variants with a seasonal fluctuating pattern than found in the original
\textcite{Bergland2014-ij} study. While surprising, this could be explained by
the presence of temporal structure across the samples not consistent with
seasonal fluctuating selection. Some fraction of the permutations happen to fit
this structure well, leading to an enrichment of small p-values. This
non-seasonal temporal structure is also evident in our temporal covariances
(Supplementary Figure XXX), where we see strong evidence of selection (non-zero
temporal covariances), yet the pattern does not follow that of seasonal
fluctuating selection.  \gc{ADD MORE HERE}

>>>>

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change ($\Delta_{_{10}} p_t$) in some initial reference generation $s$ (the row
of the matrix), and some later timepoint $t$ (the column of the matrix). We
corrected these matrices for biases created due to sampling noise, and
normalize the entries for heterozygosity (see Supplementary Material Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
the bias correction). The covariances are expected to be zero when only drift
is acting in a closed population, as only heritable variation for fitness can
create covariance between allele frequency changes \parencite{Buffalo2019-io}.
Averaging across the ten replicate temporal covariances matrices, we find
temporal covariances that are statistically significant (95\% block bootstraps
CIs do not contain zero), consistent with linked selection perturbing
genome-wide allele frequency changes over very short time periods. The
covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals, as expected when
directional selection affects linked variants' frequency trajectories until
ultimately linkage disequilibrium and additive genetic variance for fitness
decay \parencite{Buffalo2019-io}. The temporal covariances per replicate are
noisier but this general pattern holds; see Supplementary Figures
\ref{suppfig:barghi-cov-panels}.  \textcite{Barghi2019-qy}'s design means that
the covariances we see in adjacent time intervals are on average ten
generations apart, and given the temporal decay in covariance we see, the
covariances on shorter time-scales may well be higher yet (see Supplementary
Material Section \ref{supp:barghi-covs} for more details).

@@

Finally, we observed that in the \gc{longest study}  we   %run evolve-and-resequence 
analyzed \parencite{Barghi2019-qy}, some genome-wide temporal covariances
\gc{become} negative at future timepoints (see the first two rows in
Figure \ref{fig:figure-1} A). \gc{This shows that alleles that were on
average going up initially are no going down in frequency, i.e. that
the average direction of selection experienced by alleles has
flipped. This must reflect either a change in the environment or the
genetic background, due to changing epistatic relationships among
alleles or  recombination breaking up selective alleles.}
\gc{Such reversals of selective dynamics could be occuring at other time-points but} the
signal of a change in the direction of selection \gc{at particular
  loci} may be washed out when we calculate our  genome-wide \gc{average} temporal
covariances. To address this limitation, we
calculated \gc{the distribution of the} temporal covariances over
100kb genomic windows \gc{(Figure \ref{fig:figure-3}, pooling
across all replicates see Supplementary Figure \ref{suppfig:barghi-offset-replicate-panels} for individuals replicates).}
\gc{The}
covariance of each tile is noisy, \gc{due to sampling and genetic
  drift, shown in }. \gc{The neutral distribution of the covariance is
  complicated due to} linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). 
\gc{Thus} we have developed a permutation-based procedure that constructs a null
distribution by randomly flipping the signs of the    %of the
                                %genome-wide covariances as they would
                                %look if only neutral genetic drift was acting. We do this 
allele frequency changes per-genomic window. \gc{This destroys}  %%  GC bit about 1,000 times seems more like figgure caption or methods
the systematic covariances created by linked selection but creates a sampling
distribution of the covariances spuriously created by neutral genetic
drift while preserve the complex dependencies
between adjacent loci created by linkage disequilibrium (see Supplementary
Material Section \ref{supp:empirical-null} for more details on this approach).
This empirical neutral null distribution is conservative in the sense that the
variances of the covariances are wider than expected under drift
alone \gc{as they include the effect of selection on the allele
  frequency change within a time-interval, just not between
  time-intervals.} We see, Figure \ref{fig:figure-3} A and B, that
windowed temporal covariances between close timepoints are skewed positive (a
heavy right tail), while between more distant timepoints these windowed
temporal covariances tend to shift to become more negative (a heavy
left tail).
We quantified the degree to which the left and right
tails \gc{are inflated compared to the null distribution as a function
  of time}, and see excesses in both tails in Figure
\ref{fig:figure-3} C. This finding is also
robust to sign-permuting allele frequency changes on a chromosome-level, the
longest extent that gametic linkage disequilibria can extend {\bf ????} (Supplementary
Figure \ref{suppfig:barghi-tailprobs-seqid}).  \gc{We} see a striking
pattern that \gc{the windowed} covariances not only decay towards zero, but in fact become
negative through time, consistent with \gc{many} regions in the genome having had a
reversed fitness effect at later timepoints.

@@

We would like to thank the authors of the original studies we've analyzed,
including Christian Schl{\"o}tterer, John Kelly, Kimberly Hughes, Frank Chan,
Campbell Rolian, Nick Barton, Alan Bergland, and Dmitri Petrov.  We would also
like to thank Doc Edge for helpful statistical advice, and Matt Osmond for
helpful discussions.

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[]{figures/barghi-cov-panels.pdf}

  \caption{The temporal covariances from the \textcite{Barghi2019-qy} study,
  for each replicate individually.}

  \label{suppfig:barghi-cov-panels}
\end{figure}

@@

We also applied our temporal covariance approach to \textcite{Bergland2014-ij},
which found evidence of genome-wide fluctuating selection between Spring and
Fall seasons across three years \emph{Drosophila melanogaster}. As described in
\textcite{Buffalo2019-io}, we might expect positive covariances between like
seasons (i.e. Spring 2010 and Spring 2011), and negative covariances between
dislike seasons (i.e. Spring 2010 and Fall 2011) when fluctuating selection is
strong and acts genome-wide, as the original study found. However, while we
find temporal covariances that are non-zero, we find only weak support for a
seasonal fluctuating model driving these covariances. In Supplementary Figure
\ref{suppfig:bergland-covs-figure}, we show the temporal covariances from
\vb{varying reference generations, across seasonal transitions that are alike (e.g.
the covariance between the allele frequency changes between Fall 2009 and
Spring 2009, and frequency changes between Fall 2010 and Spring 2010), and
dislike (e.g. the covariance between the allele frequency change between Fall
2009 and Spring 2009, and the frequency changes between Spring 2010 and Fall
2009)}. The first row of temporal covariance matrix is consistent with
fluctuating selection operating for two timepoints, as the first covariance is
negative, and the second is positive, and later covariances are not
statistically differentiable from zero (which could occur if LD and additive
genetic variance decay). However, the all other temporal covariances do not fit
the pattern we would expect under genome-wide fluctuating selection.

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[]{figures/bergland-combined-hists.pdf}
  \caption{}
  \label{suppfig:bergland-pvalue-hist}
\end{figure}

@@

To investigate whether there is a genome-wide evidence of an enrichment of
fluctuating selection we created an empirical null distribution by randomly
permuting the season labels and re-running the per-locus seasonal GLM model, as
proposed by \parencite{Machado2018-cs}. We find, regardless of whether we
permute at the locus-level or the permutation replicate-level, that the
observed seasonal p-value distribution \textcite{Bergland2014-ij} is not
enriched for significant p-values beyond what we would expect from the
permutation null. In fact, there appears there is more enrichment for low
p-values when seasonal labels are randomly permuted (Supplementary Figure
\ref{suppfig:bergland-pvalue-hist}, suggesting by random chance we might expect
more variants with a seasonal fluctuating pattern than found in the original
\textcite{Bergland2014-ij} study. While surprising, this could be explained by
the presence of temporal structure across the samples not consistent with
seasonal fluctuating selection. Some fraction of the permutations happen to fit
this structure well, leading to an enrichment of small p-values. This
non-seasonal temporal structure is also evident in our temporal covariances
(Supplementary Figure \ref{suppfig:bergland-covs-figure}), where we see strong
evidence of selection (non-zero temporal covariances), yet the pattern does not
follow that of seasonal fluctuating selection.  \gc{ADD MORE HERE}

\begin{figure}[!ht]
  \centering
  \includegraphics[]{figures/bergland-covs-figure.pdf}

  \caption{Temporal covariances from the \textcite{Bergland2014-ij} study, from
    varying reference generations (e.g. rows along the temporal covariance
    matrix). Each covariance is labeled indicating whether the covariance is
    between two like seasonal transitions (e.g. the covariance between allele
    frequency changes from fall to spring in one year, and fall to spring in
    another or two dislike seasons (e.g. the covariance between fall to spring
    in one year, and spring to fall in another year). Covariances between like
    transitions are expected to be positive when there is a genome-wide effect
    of fluctuating selection (and these labels are colored blue), while
    covariances between dislike transitions are expected to be negative (and
    these labels are colored red). 95\% confidence intervals were constructed
    by a block-bootstrapping procedure where the blocks are megabase tiles.}

  \label{suppfig:bergland-covs-figure}
\end{figure}
